General Game Playing and Monte-Carlo Tree Search

### 2013-11-15 doc/paper/paper.tex
Style changes and other nitpicks/improvements

\section{Introduction}

\section{Critique}

@@

\subsubsection{Monte-Carlo Tree Search}

@@

\subsubsection{Summary}

\subsubsection{Critique}

@@

\subsubsection{Summary}

\subsubsection{Critique}

\subsection{Synthesis}
% Maybe I won't be able to write more than a few sentences
% In which case, remove the Synthesis section

@@

\subsubsection{Clojure}
The implementation of General Game Playing Monte-Carlo Tree Search presented in this paper was implemented in the programming language Clojure. Clojure is a Lisp dialect that can be compiled to run on the Java Virtual Machine, .NET Common Language Runtime, and JavaScript Virtual Machine. It was chosen for a variety of reasons
\begin{itemize}
% TODO: BOLD TITLE TERMS
\item \emph{Data Structures} Clojure has a rich collection of immutable, persistent data structures including lists, vectors, sets, and maps.
\item \emph{Libraries} Clojure has multiple libraries suited for the project, including pattern matching and logic programming libraries.
\item \emph{Highly Functional} Clojure has a large set of higher-order functions that can operate on any of its collections.
\item \emph{Homoiconicity} Part of this project involves translating Game Description Language into Clojure. This task would not have been possible in the timespan it was done in in a language that is not code-as-data.
\item \emph{Comfort} On a personal level, I am most comfortable programming in Clojure.
\end{itemize}
\subsubsection{core.logic}
%FIXME: CAN'T LINEBREAK AFTER THIS SUBSUBSECTION
The primary Clojure library used in the implementation was core.logic, which is a port of the logic programming language miniKanren as formulated by William Byrd in his 2010 PhD thesis \cite{byrd2010relational} and by Dan Friedman in his book The Reasoned Schemer \cite{reasonedschemer}.

@@

(Note: The (\texttt{run* [q] \textless RELATION\textgreater)} macro means 'give me all the values of q such that the relation (with q substituted in accordingly) succeeds')

@@

\begin{lstlisting}[frame=single, caption=Relation to generate all pairs of whole numbers (x y) such that x+y equals 3 or 4]
(defn pair-example [p]
  (fresh [x y]
    (== [x y] p)
    (in x y (interval 0 4))
    (conde
      [(+ x y 3)]
      [(+ x y 4)])))

@@

\subsection{GDL Translation to core.logic}
\subsubsection{Motivation}
Game Description Language is trivially derived from English. Below are the English and GDL descriptions of legality in Tic-Tac-Toe.
\begin{lstlisting}[frame=single, caption=The Legal relation for Tic-Tac-Toe expressed in English]
;It is Legal for player w to mark cell (m,n)
;if it is true that cell (m,n) is blank and
;it is true that it is player w's turn to
;move. If it is X's turn to move, O can only
;noop. If it is O's turn to move, X can only
;noop. 
\end{lstlisting}
\begin{lstlisting}[frame=single, caption=The Legal relation for Tic-Tac-Toe expressed in GDL]
(<= (legal ?w (mark ?m ?n))
    (true (cell ?m ?n b))
    (true (control ?w)))

@@

\subsubsection{Implementation}

\subsubsection{Results}
The translation algorithm was tested on four Game Description Language descriptions that can be found on the Stanford Gamemaster website: Tic-Tac-Toe, Chinkook, Chinese Checkers, and Pentago. These games are diverse in factors such as board structure, number of players, types of moves, synchronicity (??) of moves, etc (!!).

@@

\subsection{Goal Stability Early Cutoff}
Due to the modular nature of the Monte-Carlo Tree Search code, it was easy to implement the Goal Stability Early Cutoff extension \cite{finnsson2012generalized}. In the playout function, which originally just randomly plays a game to completely, a list of the result of the Goal relation is stored for each player.

@@

The implementation did not break any game that was played using it; however, none of the games that were used for testing had a fine-grained enough Goal relation for the Goal Stability Early Cutoff extension to frequently (or sometimes ever) come into effect.

@@

\section{Conclusion}

>>>>

\section{Introduction}
% TODO
\section{Critique}

@@

\subsubsection{Monte-Carlo Tree Search}
% TODO

@@

\subsubsection{Summary}
% TODO

\subsubsection{Critique}
% TODO

@@

\subsubsection{Summary}
% TODO

\subsubsection{Critique}
% TODO

\subsection{Synthesis}
% TODO
% Maybe I won't be able to write more than a few sentences
% In which case, remove the Synthesis section

@@

\subsubsection{Clojure}
The implementation of General Game Playing Monte-Carlo Tree Search presented in this paper was implemented in the programming language Clojure. Clojure is a Lisp dialect that can be compiled to run on the Java Virtual Machine, .NET Common Language Runtime, and JavaScript Virtual Machine. It was chosen for a variety of reasons
\begin{itemize}

\item \emph{Data Structures} Clojure has a rich collection of immutable, persistent data structures including lists, vectors, sets, and maps.
\item \emph{Libraries} Clojure has multiple libraries suited for the project, namely pattern matching and logic programming libraries.
\item \emph{Highly Functional} Clojure has a large set of higher-order functions that can operate on any of its collections.
\item \emph{Homoiconicity} Part of this project involves translating Game Description Language into Clojure. This task would not have been possible in the timespan it was done in in a language that is not code-as-data.
\item \emph{Comfort} On a personal level, I am most comfortable programming in Clojure.
\end{itemize}
\subsubsection{core.logic}

The primary Clojure library used in the implementation was core.logic, which is a port of the logic programming language miniKanren as formulated by William Byrd in his 2010 PhD thesis \cite{byrd2010relational} and by Dan Friedman in his book The Reasoned Schemer \cite{reasonedschemer}.

@@

(Note: The (\texttt{run* [q] \textless RELATION\textgreater)} macro means ``give me all the values of q such that the relation (with q substituted in accordingly) succeeds'')

@@

\begin{lstlisting}[frame=single, caption=Relation to generate all pairs of whole numbers (x y) such that x+y equals 3 or 4]
(defn pair-example [p]
  (fresh [x y]
    (== [x y] p)
    (in x y (interval 0 Infinity))
    (conde
      [(+ x y 3)]
      [(+ x y 4)])))

@@

\subsection{GDL Translation to core.logic}
\subsubsection{Motivation}
Game Description Language is trivially derived from English. Below are the English and GDL descriptions of legality in Tic-Tac-Toe.
\begin{lstlisting}[frame=single, caption=The Legal relation for Tic-Tac-Toe expressed in English]
;It is Legal for player w to mark cell (m,n)
;if it is true that cell (m,n) is blank and
;it is true that it is player w's turn to
;move. If it is X's turn to move, O can only
;noop. If it is O's turn to move, X can only
;noop. That is, they take turns. 
\end{lstlisting}
\begin{lstlisting}[frame=single, caption=The Legal relation for Tic-Tac-Toe expressed in GDL]
(<= (legal ?w (mark ?m ?n))
    (true (cell ?m ?n b))
    (true (control ?w)))

@@

\subsubsection{Implementation}
% TODO
% environment
% collect and transform
% eval
% methods for updating true and does relation (membero for true)

\subsubsection{Results}
% TODO: Clean this up and expand
The translation algorithm was tested on four Game Description Language descriptions that can be found on the Stanford Gamemaster website: Tic-Tac-Toe, Chinkook, Chinese Checkers, and Pentago. These games are diverse in factors such as board structure, number of players, types of moves, synchronicity (??) of moves, etc (!!).

@@

\subsubsection{Implementation}
% TODO

\subsubsection{Results}
% TODO

\subsection{Goal Stability Early Cutoff}
\subsubsection{Implementation}
Due to the modular nature of the Monte-Carlo Tree Search code, it was easy to implement the Goal Stability Early Cutoff extension \cite{finnsson2012generalized}. In the playout function, which originally just randomly plays a game to completely, a list of the result of the Goal relation is stored for each player.

@@

\subsubsection{Results}
The implementation did not break any game that was played using it; however, none of the games that were used for testing had a fine-grained enough Goal relation for the Goal Stability Early Cutoff extension to frequently (or sometimes ever) come into effect.

@@

\section{Conclusion}
% TODO
